Enzo: An Adaptive Mesh Refinement Code for Astrophysics

### 2013-06-12 introduction.tex
alexei kritsuk changes

The \enzo\ code has been extensively used over the last two decades in
a wide variety of problems, resulting in the publication of more than
100 peer-reviewed papers based on results obtained with the code. The
variety of astrophysical systems that \enzo\ has been used for include
galaxies \citep{2003ApJ...587...13T, 2012MNRAS.425..641L, 2012ApJ...749..140H}, galaxy
clusters \citep{Loken02, Xu11, Skillman13}, the interstellar medium
\citep{Slyz05}, the intergalactic medium \citep{Fang01,
2011ApJ...731....6S}, cooling flows \citep{Li12, 2013ApJ...763...38S},
turbulence \citep{Kritsuk04, 2009JPhCS.180a2020K, Collins11}, the
formation of the first stars \citep{ABN02, 2007ApJ...654...66O,
2009Sci...325..601T, Xu08}, and the formation of stars in our own Galaxy
\citep{Collins11, Collins12a}.

Numerical simulations of astronomical phenomena now play a key role,
along with observations and analytic theory, in pushing forward our
understanding of the cosmos \citep[e.g.,][]{DecadalSurvey01,
DecadalSurvey10}.  But along with this role comes responsibility.  We
believe that those developing simulation tools must fulfill two key
obligations: the first is to make those tools available to the
community as a whole, much in the way that astronomical data is now
regularly made publicly available.  The second is to document, test
and refine those methods so that they can be critically evaluated and
expanded upon by others.  This paper represents one component of our
attempt to meet these obligations with respect to the \enzo\ code.

>>>>

The \enzo\ code has been extensively used over the last two decades in
a wide variety of problems, resulting in the publication of more than
100 peer-reviewed papers based on results obtained with the code. The
variety of astrophysical systems that \enzo\ has been used for include
galaxies \citep{2003ApJ...587...13T, 2012MNRAS.425..641L, 2012ApJ...749..140H}, galaxy
clusters \citep{Loken02, Xu11, Skillman13}, the interstellar medium
\citep{Slyz05}, the intergalactic medium \citep{Fang01,
2011ApJ...731....6S}, cooling flows \citep{Li12, 2013ApJ...763...38S},
turbulence \citep{Kritsuk04, 2007ApJ...665..416K, 2009JPhCS.180a2020K, Collins11}, the
formation of the first stars \citep{ABN02, 2007ApJ...654...66O,
2009Sci...325..601T, Xu08}, and the formation of stars in our own Galaxy
\citep{Collins11, 2011ApJ...727L..20K, Collins12a}.

Numerical simulations of astronomical phenomena now play a key role,
along with observations and analytic theory, in pushing forward our
understanding of the cosmos \citep[e.g.,][]{DecadalSurvey01,
DecadalSurvey10}.  But along with this role comes responsibility.  We
believe that those developing simulation tools must fulfill two key
obligations: the first is to make those tools available to the
community as a whole, much in the way that astronomical data are now
regularly made publicly available.  The second is to document, test
and refine those methods so that they can be critically evaluated and
expanded upon by others.  Our public release of the \enzo\ code represents our
attempt to meet the first of these obligations; this paper represents
our attempt to meet the second.

### 2013-06-12 table_of_variables_and_constants.tex
alexei kritsuk changes

  \multicolumn{2}{l}{\underline{Star Formation And Feedback}} \\[5pt]
  $t_{\rm cool}$ & cooling time \\
  $t_{\rm dyn}$ & local dynamical time \\
  $t_{\rm form}$ & star particle formation time \\
  $m_b$ & baryonic mass of cell \\
  $m_J$ & Jeans mass of cell \\
  $m_*$ & star particle mass \\
  $m_{\rm *min}$ & minimum star particle mass \\
  $f_{\rm *eff}$ & star formation efficiency parameter \\
  $f_{\rm Zb}$ & metallicity fraction of baryon gas in cell \\
  $f_{\rm m*}$ & fraction of star particle's mass returned to cell \\
  $f_{\rm Z*}$ & fraction of star particle's metal mass returned to cell \\
  $f_{\rm SN}$ & fraction of stellar rest mass energy deposited as feedback \\
  $\Delta m_{\rm sf}$ & mass of stars formed in current timestep \\
  $E_{\rm add}$ & feedback energy added to cell \\
  $T_{\rm dyn,min}$ & user-defined minimum dynamical time \\
  $\eta$ & overdensity threshold for star formation \\
  $\rho_{\rm SF}$ & constant proper density threshold for star formation \\
  $\tau_*$ & characteristic star formation timescale (for Schmidt-Law method) \\[3pt]

  \multicolumn{2}{l}{\underline{Natural Constants}} \\[5pt]
  $G$ & gravitational constant \\
  $H$ & Hubble constant \\
  $k_B$ & Boltzmann constant \\
  $M_\odot$ & mass of Sun \\
  $m_H$ & mass of hydrogen atom \\
  $m_e$ & mass of electron \\
  $c$ & speed of light \\
  $c_s$ & sound speed \\
  \enddata

>>>>

  \multicolumn{2}{l}{\underline{Star Formation And Feedback}} \\[5pt]
  $t_{\rm cool}$ & cooling time \\
  $t_{\rm dyn}$ & local dynamical time \\
  $t_{\rm form}$ & star particle formation time \\
  $m_{\rm b}$ & baryonic mass of cell \\
  $m_{\rm J}$ & Jeans mass of cell \\
  $m_*$ & star particle mass \\
  $m_{\rm *min}$ & minimum star particle mass \\
  $f_{\rm *eff}$ & star formation efficiency parameter \\
  $f_{\rm Zb}$ & metallicity fraction of baryon gas in cell \\
  $f_{\rm m*}$ & fraction of star particle's mass returned to cell \\
  $f_{\rm Z*}$ & fraction of star particle's metal mass returned to cell \\
  $f_{\rm SN}$ & fraction of stellar rest mass energy deposited as feedback \\
  $\Delta m_{\rm sf}$ & mass of stars formed in current timestep \\
  $E_{\rm add}$ & feedback energy added to cell \\
  $T_{\rm dyn,min}$ & user-defined minimum dynamical time \\
  $\eta$ & overdensity threshold for star formation \\
  $\rho_{\rm SF}$ & constant proper density threshold for star formation \\
  $\tau_*$ & characteristic star formation timescale (for Schmidt-Law method) \\[3pt]

  \multicolumn{2}{l}{\underline{Natural Constants and Values}} \\[5pt]
  $G$ & gravitational constant \\
  $H$ & Hubble constant \\
  $k_B$ & Boltzmann constant \\
  $M_\odot$ & mass of Sun \\
  $m_H$ & mass of hydrogen atom \\
  $m_e$ & mass of electron \\
  $c$ & speed of light \\
  $c_{\rm s}$ & sound speed \\
  \enddata

### 2013-06-12 tests-doublemach.tex
alexei kritsuk changes

\subsubsection{Double mach reflection}
\label{sec.tests.doublemach}

The double mach reflection test is a classic two-dimensional test of
hydrodynamic algorithms originally described in
\citet{1984JCoPh..54..115W} \citep[and more recently
in][]{2008ApJS..178..137S}.  In this problem, shown in
Figure~\ref{fig.doublemach}, a shock is injected at an angle to a
reflecting surface (the -y-boundary), and a jet appears along the
reflecting surface.  The ideal solution is self-similar, and the
appearance of this solution is highly sensitive to numerical
diffusion.  If numerical noise is present, a Kelvin-Helmholz
instability develops along this jet and breaks the self-similarity.

@@

It is instructive to compare Figure~\ref{fig.doublemach} to Figure 9
in~\citet{1984JCoPh..54..115W}.  By the end of the simulation, a dense
jet is apparent at the leading edge of the shock, propagating along
the x-axis.  The shape of this jet is sensitive to numerical
diffusion, and our figure compares favorably to the higher-order
images from Woodward \& Colella.

>>>>

\subsubsection{Double Mach reflection}
\label{sec.tests.doublemach}

The double Mach reflection test is a classic two-dimensional test of
hydrodynamic algorithms originally described in
\citet{1984JCoPh..54..115W} \citep[and more recently
in][]{2008ApJS..178..137S}.  In this problem, shown in
Figure~\ref{fig.doublemach}, a shock is injected at an angle to a
reflecting surface (the -y-boundary), and a jet appears along the
reflecting surface.  The ideal solution is self-similar, and the
appearance of this solution is highly sensitive to numerical
diffusion.  If numerical noise is present, a Kelvin-Helmholz
instability develops along this jet and breaks the self-similarity.

@@

It is instructive to compare Figure~\ref{fig.doublemach} to Figure 9
in~\citet{1984JCoPh..54..115W}.  By the end of the simulation, a dense
jet is apparent at the leading edge of the shock, propagating along
the x-axis.  The shape of this jet is sensitive to numerical
diffusion, and our figure compares favorably to those shown in Woodward \& Colella.

### 2013-06-12 tests-vandv.tex
alexei kritsuk changes

Several comparisons have been made that focus specifically on
hydrodynamics solvers and fluid behavior.  In particular, the work of
\citet{2007MNRAS.380..963A} and \citet{Tasker2008} perform
direct comparisons between several grid- and particle-based
codes for a variety of fluid-centric test problems (including shocked gas clouds,
self-gravitating, translating clouds, and Sedov-Taylor explosions),
and show that \enzo\ is comparable or superior in behavior to the
other grid-based hydrodynamics codes involved in the comparison, and
provide useful information on the sort of practical challenges that a
user of an
AMR code such as \enzo\ may experience.  More specific comparisons, including one
specifically testing the linear and nonlinear growth of the
Kelvin-Helmholz instability \citep{2012ApJS..201...18M}, as well as a
comparison that more
broadly examines Galilean invariance in grid-based codes
\citep{2010MNRAS.401.2463R}, show that \enzo, and in particular its
implementation of the Piecewise Parabolic Method hydrodynamics solver,
converge to the correct solution as expected, and generally provide
less diffusive solutions than lower-order codes, including those that
use artificial viscosity.  Finally, there have been two code comparison
projects that focus on turbulence simulations.  The first studied the
behavior of decaying isothermal supersonic turbulence
\citep{2009A&A...508..541K}, and the second examined supersonic
magnetohydrodynamical turbulence \citep{2011ApJ...737...13K}.  Both
included the \enzo\ code, with the former testing the PPM
hydrodynamics and the latter the constrained transport MHD
implementation of \citet{Collins10}.  In both cases, \enzo\ performed similarly to other
grid-based codes that use Godunov-based fluid solvers, and typically
had better effective resolution than particle-based codes when using the
same number of particles as the number of grid cells in the \enzo\ simulation.

>>>>

Several comparisons have been made that focus specifically on
hydrodynamics solvers and fluid behavior.  In particular, the work of
\citet{2007MNRAS.380..963A} and \citet{Tasker2008} perform
direct comparisons between several grid- and particle-based
codes for a variety of fluid-centric test problems (including shocked gas clouds,
self-gravitating, translating clouds, and Sedov-Taylor explosions),
and show that \enzo\ is comparable or superior in behavior to the
other grid-based hydrodynamics codes involved in the comparison, and
provide useful information on the sort of practical challenges that a
user of an
AMR code such as \enzo\ may experience.  More specific comparisons, including one
specifically testing the linear and nonlinear growth of the
Kelvin-Helmholz instability \citep{2012ApJS..201...18M}, as well as a
comparison that more
broadly examines Galilean invariance in grid-based codes
\citep{2010MNRAS.401.2463R}, show that \enzo, and in particular its
implementation of the Piecewise Parabolic Method hydrodynamics solver,
converge to the correct solution as expected, and generally provide
less diffusive solutions than lower-order codes, including those that
use artificial viscosity.  Finally, there have been two code comparison
projects that focus on turbulence simulations.  The first studied the
behavior of decaying isothermal supersonic turbulence
\citep{2009A&A...508..541K}, and the second examined supersonic
magnetohydrodynamical turbulence \citep{2011ApJ...737...13K}.  Both
included the \enzo\ code, with the former testing the PPM
hydrodynamics and the latter both the constrained transport MHD
implementation of \citet{Collins10} and the Dedner method of \citet{WangAbelZhang08}.  In both cases, \enzo\ performed similarly to other
grid-based codes that use Godunov-based fluid solvers, and typically
had better effective resolution than particle-based codes when using the
same number of particles as the number of grid cells in the \enzo\ simulation.
